---
title: "advertising"
output: html_document
date: "2025-02-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Correlation Test

In the *advertising* dataset, test whether there is a *positive* relationship between the target variable “`approved`” and the variable “`spend`” with $\alpha=0.025$. 

For the correlation test, we use the function `cor.test` as follows

```{r}
cor_test<- cor.test(x = advertising$spend, 
          y = advertising$approved, 
          alternative = "greater", 
          conf.level = 0.975)
```

a. **Report the hypotheses $H_0$ and $H_1$. Report the value of *p*-value. Explain you accept or reject the null hypothesis, at $\alpha=0.025$.** </br>
Null and Alternative Hypotheses: </br>

  \[
  H_0: \rho \leq 0
  \]

  \[
  H_1: \rho > 0
  \]

p-value: `r cor_test$p.value` </br>
Correlation Coefficient: `r round(cor_test$estimate, 4)` </br>
Confidence Interval: `r round(cor_test$conf.int[1], 4)` to `r round(cor_test$conf.int[2], 4)` at the 97.5% confidence level

Since the p-value is significantly less than the significance level of α=0.025, we reject the null hypothesis. This indicates strong evidence of a positive relationship between the variables "spend" and "approved". Therefore, there is a statistically significant positive correlation between how much is spent on advertising and the approval rate.

b. **What conclusions can be drawn regarding "a *positive* relationship between the two variables".** </br>
A positive relationship between the two variables implies that increasing the advertising spend is associated with higher approval rates, which could inform future marketing strategies.